master thesis

### 2018-03-04 Paper - Latex/Paper - Skizze - v2.tex
Results - Multifurcation; BICs added

      %---------------------------------------------------------------------------------------------------
      \subsubsection{Data artifacts}
        Zu der Frage, warum gibt es $nrChildren==1$. Es gibt tatsächlich 55700 Knoten mit nur einem Kind. 
          Davon ist der großteil direkt vor einem Blatt (51744), aber 3956 ($height >2$) sind irgendwo im 
          Baum. Ein paar Beispiele: \\
        Nephroselmidophyceae: (class) \\
        https://tree.opentreeoflife.org/opentree/argus/ottol@1038762 \\
        Phrynocrinidae: (family) \\
        https://tree.opentreeoflife.org/opentree/argus/ottol@3647979 \\
        Elaeocarpus sylvestris: \\
        https://tree.opentreeoflife.org/opentree/argus/opentree9.1@ott166969 \\
        Was bedeutet das für unsere Analyse? Ignorieren wir die $\sim$ 4000 internen und Zählen sie zu den 
          Blättern, oder zählen wir alle zu den Internen oder puzzlen wir sie einzeln raus. D.h. nur die 
          mit der Zusatzeigenschaft ($height == 2$) sind auch Blätter. \\
        Es könnte aber auch Fälle geben wie: ...-O-O-O-O-O, dann könnte man auch einen Knoten mit höherer 
          Höhe als Blatt zählen. Das müsste dann gelten $min height = max height$. \\ 
        \begin{lstlisting}[gobble=8]
          leaf.taxa <-  all.taxa[all.taxa$nr_children==0, ]
          extendedLeaf.taxa <- all.taxa[all.taxa$nr_children<2 & min.height==max.height, ]
          inner.taxa <- all.taxa[all.taxa$nr_children>1 | (all.taxa$nr_children==1 & min.height!=max.height), ]
        \end{lstlisting}
        \todo{macht extendedLeaf wirklich einen Sinn?}

@@

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Multifurcation}
      One property of the tree is its ridge of multifurcation. A complete kinship tree would be binary, 
        but since we only work with a synthesis tree of individual trees, this tree is multifurcated. 
        For a first overview we collected for every node its number of children (degree $-1$), and 
        plottet this in an histogram. In Figure \ref{fig:childrenOfNodes} you can see how much nodes we 
        have for every degree. \\
      \todotext{Subfigure \ref{fig:childrenOfNodes_innerNodes} beschränkt sich hierbei nur auf die inner 
        nodes während Subfigure \ref{fig:childrenOfNodes_allNodes} alle Knoten abbildet.} \\
      \todo{Welchen der beiden Plots wollen wir nehmen, einer genügt eigentlich:}
      \begin{figure}[h!]
        \begin{subfigure}{.5\textwidth}
          \centering
          \includegraphics[width=.8\linewidth]{Figures/multifurc.pdf}
          \caption{all nodes}
          \label{fig:childrenOfNodes_allNodes}
        \end{subfigure}%
        \begin{subfigure}{.5\textwidth}
          \centering
          \includegraphics[width=.8\linewidth]{Figures/nr_children.pdf}
          \caption{inner nodes (shiftet to 0)}
          \label{fig:childrenOfNodes_innerNodes}
        \end{subfigure}
        \caption{plots of children of nodes}
        \label{fig:childrenOfNodes}
      \end{figure}
      As you can see, we are very far from a binary tree. Therefore, the impact of multifurcations on 
        our research question has been investigated:
      \begin{itemize}
        \item stärke der Multifurcation in verschiedenen Subbäumen
        \item zusammenhang zur tiefe im Baum
        \item zusammenhang zu höhe im baum (max, min, mean)
        \item einfluss auf die vorhersage des Castor und Fitch Algorithmus (Simulation) \todo{...}
      \end{itemize}

      "Nested models were compared using likelihood ratio tests, models using different predictors were 
      compared according to their deviance and AIC."

@@

%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
\chapter{Results}
  Ein großer Punkt in diesem Kapitel ist das Ergebnis der Untersuchung der Datenlage. Wie ist der 
    Stand? Welchen Einfluss hat das auf unser eigentliches Ergebnis? Was kann man dagegen tun? \\
  Ansonsten geht es in diesem Kapitel hauptsächlich um die tatsächliche Rekonstruktion der states.
    Dass heißt zum einen der der inneren Knoten und dadurch der Untersuchung von Origins und Losses
    und zum anderen der Vorhersage von unbekannten states von Leaf nodes. \\
  Und natürlich dem Ergebnis der Simulation.

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Metadata analysis}

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Taxa}
      Die Untersuchung der Taxa der Eukaryoten hat das folgende ergeben:
      Es gibt 3 Kingdoms: Chloroplastida, Metazoa, Fungi. \\
      \todo{Hier fehlt nichts. Ref einfügen.} \\
      Phyla (53): \\
      Acanthocephala, Amoebozoa, Apicomplexa, Arthropoda, Ascomycota, Bacillariophyta, Basidiomycota, 
        Brachiopoda, Bryozoa, Chaetognatha, Chlorophyta, Chordata, Chromerida, Chytridiomycota, 
        Ciliophora, Cnidaria, Colponemidia, Ctenophora, Cycliophora, Echinodermata, Entoprocta, 
        Entorrhizomycota, Euglenida, Foraminifera, Gastrotricha, Glomeromycota, Gnathostomulida, 
        Haplosporida, Haptophyta, Hemichordata, Kinorhyncha, Loricifera, Microsporidia, Mollusca, 
        Myzostomida, Nematoda, Nematomorpha, Nemertea, Onychophora, Orthonectida, Phaeophyceae, 
        Picozoa, Placozoa, Platyhelminthes, Porifera, Priapulida, Rhodophyta, Rhombozoa, Rotifera, 
        Streptophyta, Tardigrada, Xanthophyceae \\
      Wobei von Streptophyta -> Anthocerotophyta, Marchantiophyta, Bryophyta, Tracheophyta als
        Phylum im Phylum gefunden und nicht einbezogen wurden und Magnoliophyta als Phylum in 
        Tracheophyta ebenfalls nicht. \\
      Außerdem haben wir 195 Klassen und 924 Ordnungen feststellen können.

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \section{Multifurcation}

       %---------------------------------------------------------------------------------------------------
       %---------------------------------------------------------------------------------------------------
       \subsection{Poisson regression}

        \begin{lstlisting}[gobble=8]
          glm(formula = multifurc ~ 1, family = "poisson", data = inner.taxa)
        \end{lstlisting}
        The intercept is $2.821 > 0$ $\Rightarrow$ there is a multifunction.
        (Intercept: Stärke der Multifurcation) \\
        Comparing the different kingdoms, we find that multifunctionality is greater in Fungi than in 
        Chloroplastida than in Metazoa:
        $$4.0999 (Fungi Intercept) > -0.9132 (Chloroplastida Intercept) > -1.4320 (Metazoa Intercept)$$
        % Degrees of Freedom: 96010 Total (i.e. Null);  96010 Residual
        Wir haben außerdem drei komplexitätsstärken von Modellen verglichen bezüglich der höhe und tiefe 
          des Baums mit dem folgenden Deviance Table:

@@

        \todotext{Nur ein Ausschnitt hier und dann... You can find the whole Deviance Table in the 
          Appendix...} \\
        Interpretation: Die Multifurkation ist sehr ungleich verteilt. Daher ist die vorhersage umso 
          genauer umso kleinere Subtrees wir betrachten. ...
        \begin{lstlisting}[gobble=8]
          Model 1: multifurc ~ kingdom
          Model 2: multifurc ~ phylum
          Model 3: multifurc ~ class
          Model 4: multifurc ~ order
            Resid. Df Resid. Dev  Df Deviance  Pr(>Chi)
          1    167222    7774454
          2    167173    7435700  49   338754 < 2.2e-16 ***
          3    167039    7337241 134    98459 < 2.2e-16 ***
          4    166352    7076068 687   261172 < 2.2e-16 ***
        \end{lstlisting}
        \anmerkung{different parameters for different taxa? (Emanuel)} \\ \\

@@

%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
\chapter{Appendices}
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Distribution of data in the taxa}\label{sec:taxaTable1}
    \begin{center}
      \begin{tabular}{ |l|r|l|r|r| }
        \hline
        Kingdom (3) & Number of Nodes & Phylum (25) & Number of Nodes & max max height \\
        \hline \hline
        Metazoa & 1 465 207 & Arthropoda & 1 170 539 & 54 \\
        & & Chordata & 106 650 & 74 \\
        & & Mollusca & 80 022 & 22 \\
        & & Platyhelminthes & 27 141 & 9 \\
        & & Nematoda & 24 564 & 23 \\
        & & Cnidaria & 14 878 & 36 \\
        & & Porifera & 11 737 & 26 \\
        & & Echinodermata & 10 654 & 14 \\
        & & Bryozoa & 8 631 & 11 \\
        & & Rotifera & 3 093 & 7 \\
        & & Nemertea & 1 793 & 7 \\
        & & Tardigrada & 1 654 & 7 \\
        & & Acanthocephala & 1 596 & 6 \\
        & & Brachiopoda & 1 055 & 9 \\
        & & Nematomorpha & 633 & 7 \\
        & & Chaetognatha & 360 & 7 \\
        & & Hemichordata & 196 & 5 \\ 
        & & Cycliophora & 11 & 5 \\ 
        % 1465207-(1170539+106650+80022+27141+24564+14878+11737+10654+8631+3093+1793+1654+1596+1055+633+360+196+11) = 0          \hline
        Fungi & 254 871 & Ascomycota & 157 045 & 19 \\ 
        & & Basidiomycota & 92 931 & 18 \\
        & & Microsporidia & 1 949 & 6 \\ 
        & & Glomeromycota & 1 490 & 6 \\
        & & Chytridiomycota & 1 456 & 6 \\
        % 254871-157045-92931-1949-1490-1456 = 0
        \hline
        Chloroplastida & 121 239 & Streptophyta & 120 731 & 49 \\
        & & Chlorophyta & 508 & 6 \\
        % 121239-120731-508 = 0
        \hline  
      \end{tabular}
    \end{center}

>>>>

@@

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Multifurcation}
      We have a very high multifurcation rate \pageref{sec:ResultsMultifurcation}.

      Therefore, the impact of multifurcations on our research question has been investigated:
      \begin{itemize}
        \item stärke der Multifurcation in verschiedenen Subbäumen
        \item zusammenhang zur tiefe im Baum
        \item zusammenhang zu höhe im baum (max, min, mean)
        \item einfluss auf die vorhersage des Castor und Fitch Algorithmus (Simulation) \todo{...}
      \end{itemize}

      "Nested models were compared using likelihood ratio tests, models using different predictors were 
      compared according to their deviance and AIC."
      \begin{lstlisting}[gobble=8]
        glm(formula = multifurc ~ 1, family = "poisson", data = inner.taxa)
      \end{lstlisting}

@@

%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
\chapter{Results}
  A big point in this chapter is the result of examining the input data. How is the situation? What 
    influence does that have on our actual result? What can we do about it? Our simulation gave us 
    some results to this. \\
  Otherwise, this chapter is mainly about the actual reconstruction of the states. This means, on 
    one hand investigation of origins and losses of the inner nodes and on the other, the prediction 
    of unknown states of leaf nodes. \\
   
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Metadata analysis}

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Taxa}
      The investigation of the taxonomy revealed that our tree has three kingdoms: Chloroplastida, 
        Metazoa, Fungi, 53 phyla, 195 classes and 924 orders. \\
      Since the analysis of the tree is not part of this work, it should be mentioned here that, 
        according to recent findings, this is not complete and we lack some taxa in every rank. For 
        example, Hans says that one distinguishes between seven and nine kingdoms 
        \cite{CavalierSmith1981}. \\
      In section \pageref{sec:listPhyla} of the appendix you can find a list of all phyla.

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \section{Multifurcation}\label{sec:ResultsMultifurcation}
      One property of the tree is its ridge of multifurcation. A complete phylogenetic tree would be 
        binary, but since we only work with a synthesis tree, this tree is multifurcated. \\
      For a first overview we collected for every node its number of children (degree $-1$), and plotted
        this in an histogram, see figure \ref{fig:childrenOfNodes}. The multifurcation affects only the
        internal nodes. We collected the number of children $-2$ of every node (a node with two children 
        is binary). 
      \begin{wrapfigure}{r}{0.6\textwidth}
        \begin{center}
          \includegraphics[trim = 0mm 0mm 30mm 0mm, clip, width=0.58\textwidth]{Figures/multifurc.pdf}
        \end{center}
        \caption{Histogram about the multifurcation of the internal nodes of the synthesis tree. \\ 
          light blue: binary; dark blue: multifunction}
        \label{fig:childrenOfNodes}
      \end{wrapfigure}
      That means it discribes the number of nodes which we have lost from the real (binary) 
      phylogenetic tree. \\
      As you can see, we are very far from a binary tree.

      %---------------------------------------------------------------------------------------------------
      \subsubsection{Data artifacts}
      At this point we also found out that there are some nodes with only one child node (55700 nodes). \\
      These are both, the most nodes are right in front of a leaf, as well as some nodes are deep in the 
        tree (3956 with height $>2$). They are probably a result from the fact that taxonomic information 
        has been incorporated into a phylogeny. \\

      Some examples:
      \begin{itemize}
        \item Nephroselmidophyceae: (class) \\
          https://tree.opentreeoflife.org/opentree/argus/ottol@1038762
        \item Phrynocrinidae: (family) \\
          https://tree.opentreeoflife.org/opentree/argus/ottol@3647979
        \item Elaeocarpus sylvestris: \\
          https://tree.opentreeoflife.org/opentree/argus/opentree9.1@ott166969
      \end{itemize}

      %---------------------------------------------------------------------------------------------------
      %---------------------------------------------------------------------------------------------------
      \subsection{Poisson regression}
        glm (generalized linear model) analysis: \\

        The intercept is $2.821 > 0$ $\Rightarrow$ there is a multifunction.
        (Intercept: Stärke der Multifurcation) \\
        Comparing the different kingdoms, we find that multifunctionality is greater in Fungi than in 
        Chloroplastida than in Metazoa:
        $$4.0999 (Fungi Intercept) > -0.9132 (Chloroplastida Intercept) > -1.4320 (Metazoa Intercept)$$
        % Degrees of Freedom: 96010 Total (i.e. Null);  96010 Residual
        Wir haben außerdem drei komplexitätsstärken von Modellen verglichen bezüglich der höhe und tiefe 
          des Baums mit dem folgenden Deviance Table:

@@

        \todotext{Nur ein Ausschnitt hier und dann... You can find the whole Deviance Table in the 
          Appendix...} \\
        Interpretation: Die Multifurkation ist sehr ungleich verteilt. Daher ist die vorhersage umso 
          genauer umso kleinere Subtrees wir betrachten. ...

        \anmerkung{different parameters for different taxa? (Emanuel)} \\ \\

        Because of the difference in the complexity of the models, we compared their BICs:
        \begin{table}
          \begin{center}
            \begin{tabular}{ |l|r| }
              \hline
              Model & BIC \\
              \hline \hline
              multifurc $\sim$ 1 & 8623935 \\
              multifurc $\sim$ kingdom & 8273333 \\
              multifurc $\sim$ phylum & 7937828 \\
              multifurc $\sim$ class & 7842157 \\
              \rowcolor{green!50}multifurc $\sim$ order & 7644249\\
              \hline \hline
              multifurc $\sim$ phylum + depth & 7934322 \\
              multifurc $\sim$ phylum + max.height & 7749121 \\
              multifurc $\sim$ phylum + min.height & 7875521 \\
              \rowcolor{green!50} multifurc $\sim$ phylum + mean.height & 7644249 \\
              \hline
              multifurc ~ phylum * depth & 7836755 \\
              multifurc ~ phylum * max.height & 7693555 \\
              multifurc ~ phylum * min.height & 7808940 \\
              \rowcolor{green!50} multifurc ~ phylum * mean.height & 7536296 \\
              \hline \hline
              multifurc $\sim$ class + depth & \\
              multifurc $\sim$ class + max.height & \\
              multifurc $\sim$ class + min.height & \\
              multifurc $\sim$ class + mean.height & \\
              \hline
              multifurc ~ class * depth & \\
              multifurc ~ class * max.height & \\
              multifurc ~ class * min.height & \\
              multifurc ~ class * mean.height &\\
              \hline
            \end{tabular} 
          \end{center}
          \caption{BIC...}
          \label{table:...} 
        \end{table}

@@

%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
\chapter{Appendices}
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{List of all phyla}\label{sec:listPhyla}

  Phyla (53): \\
  Acanthocephala, Amoebozoa, Apicomplexa, Arthropoda, Ascomycota, Bacillariophyta, Basidiomycota, 
    Brachiopoda, Bryozoa, Chaetognatha, Chlorophyta, Chordata, Chromerida, Chytridiomycota, 
    Ciliophora, Cnidaria, Colponemidia, Ctenophora, Cycliophora, Echinodermata, Entoprocta, 
    Entorrhizomycota, Euglenida, Foraminifera, Gastrotricha, Glomeromycota, Gnathostomulida, 
    Haplosporida, Haptophyta, Hemichordata, Kinorhyncha, Loricifera, Microsporidia, Mollusca, 
    Myzostomida, Nematoda, Nematomorpha, Nemertea, Onychophora, Orthonectida, Phaeophyceae, 
    Picozoa, Placozoa, Platyhelminthes, Porifera, Priapulida, Rhodophyta, Rhombozoa, Rotifera, 
    Streptophyta, Tardigrada, Xanthophyceae \\
  Wobei von Streptophyta -> Anthocerotophyta, Marchantiophyta, Bryophyta, Tracheophyta als
    Phylum im Phylum gefunden und nicht einbezogen wurden und Magnoliophyta als Phylum in 
    Tracheophyta ebenfalls nicht. \\

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Distribution of data in the taxa}\label{sec:taxaTable1}
    \begin{center}
      \begin{tabular}{ |l|r|l|r|r| }
        \hline
        Kingdom (3) & Number of Nodes & Phylum (25) & Number of Nodes & max max height \\
        \hline \hline
        Metazoa & 1 465 207 & Arthropoda & 1 170 539 & 54 \\
        & & Chordata & 106 650 & 74 \\
        & & Mollusca & 80 022 & 22 \\
        & & Platyhelminthes & 27 141 & 9 \\
        & & Nematoda & 24 564 & 23 \\
        & & Cnidaria & 14 878 & 36 \\
        & & Porifera & 11 737 & 26 \\
        & & Echinodermata & 10 654 & 14 \\
        & & Bryozoa & 8 631 & 11 \\
        & & Rotifera & 3 093 & 7 \\
        & & Nemertea & 1 793 & 7 \\
        & & Tardigrada & 1 654 & 7 \\
        & & Acanthocephala & 1 596 & 6 \\
        & & Brachiopoda & 1 055 & 9 \\
        & & Nematomorpha & 633 & 7 \\
        & & Chaetognatha & 360 & 7 \\
        & & Hemichordata & 196 & 5 \\ 
        & & Cycliophora & 11 & 5 \\ 
        % 1465207-(1170539+106650+80022+27141+24564+14878+11737+10654+8631+3093+1793+1654+1596+1055+633+360+196+11) = 0          \hline
        Fungi & 254 871 & Ascomycota & 157 045 & 19 \\ 
        & & Basidiomycota & 92 931 & 18 \\
        & & Microsporidia & 1 949 & 6 \\ 
        & & Glomeromycota & 1 490 & 6 \\
        & & Chytridiomycota & 1 456 & 6 \\
        % 254871-157045-92931-1949-1490-1456 = 0
        \hline
        Chloroplastida & 121 239 & Streptophyta & 120 731 & 49 \\
        & & Chlorophyta & 508 & 6 \\
        % 121239-120731-508 = 0
        \hline  
      \end{tabular}
    \end{center}
